universities

### 2020-08-10 inst/f1points.Rmd
include Jenkins

Formula 1 racing is an important and prestigious motor sport
\citep{test}.  Season ranking is based on a points allocation
system wherein competitors are awarded points based on race finishing
order; points accumulate additively.  The overall competition winner
is the competitor who accumulates the most points after the final
race.

@@

and a slight generalization of this would allow the incorporation of
non-finishers (DNF etc).  If, say, competitors 4 and 5 did not finish,
we would have

@@

Finding the maximum likelihood estimate for the players' strengths is
straightforward.  The `hyper2` package includes a suite of numerical
optimization routines, and because they have access to derivatives,
convergence is rapid.  A graphical illustration of the strengths is
given in figure \ref{piechartstrength}

@@

We see in figure \ref{piechartstrength} that the driver with the
largest estimated strength was Rosberg in 2016 at about 30\%, and in
years 2017-8-9 was Hamilton (at about 29\%, 37\%, and 42\%
respectively).  As an illustration of the value of likelihood methods
(as opposed to points-based methods), a likelihood ratio test
[`samep.test()`, supplied with the package] rejects
the null that Hamilton and Vettel have the same strength in
2018 ($H_0\colon \pham= \pvet$) with a likelihood ratio of
$e^{2.76}\simeq 15.8$, corresponding (by Wilks's theorem) to an
asymptotic $p$-value of about $0.02$.

>>>>

Formula 1 racing is an important and prestigious motor sport
\citep{codling2017,jenkins2010}.  Season ranking is based on a points
allocation system wherein competitors are awarded points based on race
finishing order; points accumulate additively.  The overall
competition winner is the competitor who accumulates the most points
after the final race.

@@

and this would be a forward ranking Plackett-Luce model in the
terminology of \cite{johnson2020}.  A slight generalization allows the
incorporation of non-finishers (DNF etc).  If, say, competitors 4 and
5 did not finish, we would have

@@

Finding the maximum likelihood estimate for the players' strengths is
straightforward.  The `hyper2` package includes a suite of numerical
optimization routines, and because they have access to derivatives,
convergence is rapid.  A graphical diagram of the strengths is given
in figure \ref{piechartstrength}

@@

We see in figure \ref{piechartstrength} that the driver with the
largest estimated strength was Rosberg in 2016 at about 30\%, and in
years 2017-8-9 was Hamilton (at about 29\%, 37\%, and 42\%
respectively).  As an illustration of the value of likelihood methods
(as opposed to points-based methods), a likelihood ratio test
[`samep.test()`, supplied with the package] rejects the null that
Hamilton and Vettel have the same strength in 2018 ($H_0\colon \pham=
\pvet$) with a likelihood ratio of $e^{2.76}\simeq 15.8$,
corresponding (by Wilks's theorem) to an asymptotic $p$-value of about
$0.02$.  We may also use the reified entity concept to test the null
hypothesis that Hamilton's strength remains constant 2016, where
Rosberg had the highest estimated strength, and 2017-9, where Hamilton
did; we fail to reject this null.

```{r,label=combineallyearssetup,echo=FALSE,cache=TRUE}
H2016a <- hyper2(brackets(H2016),d=powers(H2016),pnames=sub("Hamilton","Hamilton_pre", pnames(H2016)))
H2017a <- hyper2(brackets(H2017),d=powers(H2017),pnames=sub("Hamilton","Hamilton_post",pnames(H2017)))
H2018a <- hyper2(brackets(H2018),d=powers(H2018),pnames=sub("Hamilton","Hamilton_post",pnames(H2018)))
H2019a <- hyper2(brackets(H2019),d=powers(H2019),pnames=sub("Hamilton","Hamilton_post",pnames(H2019)))
Ha <- H2016a
Ha <- combine(Ha,H2017a)
Ha <- combine(Ha,H2018a) 
Ha <- combine(Ha,H2019a)
ignorethis <- samep.test(Ha,c("Hamilton_pre","Hamilton_post"))
```
